---
author: "Equipo Edumer"
editor: visual
---

# Results

## Descriptive

```{r packages, echo=FALSE}
pacman::p_load(tidyverse,
               ggpubr, 
               lme4, 
               ordinal,
               sjPlot, 
               texreg,
               kableExtra,
               easystats)


options(scipen=999)
rm(list = ls())
```

```{r data, echo=FALSE}

load("input/data/data_index.RData")
load("input/data/deservingness.RData")
load("input/data/soc_effort.RData")
load("input/data/soc_talent.RData")
load("input/data/school_effort.RData")
load("input/data/school_talent.RData")
load("input/data/data_rec.RData")

```

```{r functions, echo=FALSE}

odds_fun <- function(modelos) {
   or <- texreg::extract(modelos)
   or@coef <- exp(or@coef)
   return(or)
}

miles <- function(x) {
  format(round(as.numeric(x),0), big.mark = ".")
}

decimales <- function(x) {
  format(round(as.numeric(x), 2), decimal.mark = ",")
}

# set theme


theme_set(theme_bw())

```

@fig-marketjustice illustrates the distribution of responses to the three items on market justice preferences (healthcare, pensions, and education), indicating that most respondents tend to disagree or strongly disagree with the idea that it is just for people with higher incomes to have access to better services in these areas. The strongest opposition is observed in the healthcare sector, as the majority of respondents (66.4%) are against the idea that higher-income individuals should have access to better healthcare services. Similarly, regarding the concept of justice related to access to better pensions based on income, the level of disagreement reaches 63.0%. This decreases significantly in the field of education, as the percentage of respondents who disagree or strongly disagree with those with higher incomes having access to better education decreases to 58.8%. Only a small proportion agree that access to healthcare should be conditional on income (23.6%), and a minority strongly agree (10%), as is the case with pensions. However, this increases in the case of education, being the area where there is a significant proportion of respondents who agree or strongly agree with access based on market justice (41.2%).

This contrasts with the pattern observed for these items in the Chilean adult population (see @fig-elsoc). While both age groups tend to disagree that higher-income individuals should access better health, pension, and education services, this tendency is more pronounced in adults than in students, who show greater agreement with market justice preferences in these areas. Additionally, the hierarchy of these domains differs between groups; adults justify market justice least in education, followed by health, and then pensions, whereas students justify market principles less in health, then pensions, and lastly in education.

```{r echo=FALSE}
#| fig-cap: "Market justice preferences in health, pensions and education"
#| fig-asp: 0.6
#| label: fig-marketjustice


data_rec %>% 
  select(just_salud, just_pension, just_educ) %>% 
  sjPlot::plot_stackfrq(geom.size = 0.5,
                        expand.grid =TRUE,
                      geom.colors = "PuBu",
                      show.total = FALSE, 
                      legend.labels = c("Strongly disagree", "Disagree", "Agree", "Strongly agree"),
                      axis.labels = c("Health", "Pensions", "Education")) +
 # theme_light(base_size = 16) +
  theme(legend.position="bottom", text = element_text(size = 14),
        plot.caption = element_text(hjust = 1)) +
       labs(caption = paste0("Source: Authors calculation based on First Study of Civic Education 2017",
                             " (N = ",dim(data_rec)[1],")"))

```

Regarding the perception of meritocracy, @fig-meritocracy displays the frequency distribution of five items related to the dimensions of school and society. At the school level, there is a strong belief that both effort and talent are important for achieving good grades, with especially high agreement for effort. Specifically, while 78.6% agree or strongly agree on the importance of talent for obtaining good grades, this proportion rises to 96.8% for effort, indicating that respondents value effort more than talent in this context.

```{r echo=FALSE}
#| fig-cap: "Social and school meritocracy"
#| fig-asp: 0.8
#| label: fig-meritocracy


plot<-data_rec %>% 
  select(merito_soc, esfuerzo_soc, inteligencia_soc, esfuerzo_esc, inteligencia_esc) %>% 
  sjPlot::plot_stackfrq(geom.size = 0.5,
                        expand.grid =TRUE,
                      geom.colors = "PuBu",
                      show.total = FALSE, 
                      vjust=c(rep(c("center"),12), "bottom", "top", rep(c("center"), 6)),
                      legend.labels = c("Strongly disagree", "Disagree", "Agree", "Strongly agree"),
                      axis.labels = c("Deservingness", "Social effort", "Social talent", "School effort", "School talent")) +
  #theme_light(base_size = 16) +
  theme(legend.position="bottom", text = element_text(size = 14),
        plot.caption = element_text(hjust = 1))


plot + 
  annotate("text", x = 5.5, y = 0, label = "School meritocracy", size = 5, hjust = 0, vjust = 1) +
  annotate("text", x = 3.5, y = 0, label = "Social meritocracy", size = 5, hjust = 0, vjust = 1) +
       labs(caption = paste0("Source: Authors calculation based on First Study of Civic Education 2017",
                             " (N = ",dim(data_rec)[1],")"))
```


At the societal level, @fig-meritocracy shows that a majority of respondents agree or strongly agree that individuals are rewarded for their talents (63.7%) and efforts (62%). Nevertheless, a notable proportion perceives that both effort and talent are not adequately rewarded in society, with 38.1% and 36.3% respectively. The perception of whether people get what they deserve in society is more divided; while a slim majority (56.5% combining agree and strongly agree) agrees, a significant minority (43.6%) disagrees, indicating that this is the most contentious item among respondents. There is general agreement that people are rewarded for talent and effort in society, but the consensus is weaker compared to school meritocracy.

*Market justice preferences*

```{r echo=FALSE}
#| fig-cap: "Market justice preferences in education, health and pensions"
#| fig-asp: 1.22
#| label: fig-marketjustice

data_rec %>% 
  select(just_salud, just_educ, just_pension) %>% 
  sjPlot::plot_stackfrq(expand.grid =TRUE,
                      geom.colors = "PuBu",
                      show.total = FALSE, 
                      legend.labels = c("Strongly disagree", "Disagree", "Agree", "Strongly agree"),
                      axis.labels = c("Health", "Education", "Pensions")) +
  theme_light(base_size = 16) +
  theme(legend.position="bottom", text = element_text(size = 14),
        plot.caption = element_text(hjust = 1))
```

@fig-marketjustice shows the frequency distribution for the three market justice preference items (pensions, education, and health). The majority of respondents disagree (63.0% in total) with the idea that it is just for higher-income individuals to access better pensions. However, a notable portion (27.0%) agrees, and a smaller minority (10.1%) strongly agrees. Similar to pensions, a majority of respondents (58.8%) disagree with the idea that higher-income individuals should have access to better education. However, a considerable portion (28.6%) agrees, and a slightly larger minority (12.6%) strongly agrees compared to pensions. In the context of health, an even larger majority of respondents (66.4%) disagree with the notion that higher-income individuals should access better healthcare services. A smaller percentage (23.6%) agrees, and a minority (10.0%) strongly agrees. Overall, @fig-marketjustice shows a predominant sentiment against the justice of higher-income individuals accessing better services across all three areas. The strongest opposition is observed in the health sector. Despite the overall disagreement, a significant portion of respondents agree with the statement, particularly in education, where agreement and strong agreement are slightly higher compared to pensions and health.

*School and social meritocracy*

```{r echo=FALSE}
#| fig-cap: "Social and school meritocracy"
#| fig-asp: 1.22
#| label: fig-meritocracy

plot<-data_rec %>% 
  select(merito_soc, esfuerzo_soc, inteligencia_soc, esfuerzo_esc, inteligencia_esc) %>% 
  sjPlot::plot_stackfrq(expand.grid =TRUE,
                      geom.colors = "PuBu",
                      show.total = FALSE, 
                      vjust=c(rep(c("center"),12), "bottom", "top", rep(c("center"), 6)),
                      legend.labels = c("Strongly disagree", "Disagree", "Agree", "Strongly agree"),
                      axis.labels = c("Deservingness", "Social effort", "Social talent", "School effort", "School talent")) +
  theme_light(base_size = 16) +
  theme(legend.position="bottom", text = element_text(size = 14),
        plot.caption = element_text(hjust = 1))

plot + 
  annotate("text", x = 5.5, y = 0, label = "School meritocracy", size = 5, hjust = 0, vjust = 1) +
  annotate("text", x = 3.5, y = 0, label = "Social meritocracy", size = 5, hjust = 0, vjust = 1)
```

@fig-meritocracy shows the frequency distribution for the five meritocracy items for the dimensions of school and social meritocracy. While a significant portion of respondents (29.2%) strongly agrees that talent is important to get good grades, a larger majority (78.6% combining agree and strongly agree) supports this notion. However, the strong agreement is not as pronounced as for school effort. An overwhelming majority (73.1%) agree that students effort is important to get good grades, which is a much higher percentage compared to the strong agreement for talent (29.2%). This indicates that people believe effort is more important than talent at school. There is a strong belief that both effort and talent are important, with a particularly high agreement for effort (73.1% strongly agree).

In relation with social meritocracy, @fig-meritocracy shows that a majority of respondents (63.7% combining agree and strongly agree) agree that people are rewarded for their talent in society, with a notable portion (17.1%) strongly agreeing. A majority (62.0% combining agree and strongly agree) agree that people are rewarded for their effort in society. Similar to social talent, the perception is positive but less overwhelming than school meritocracy. The perception of whether people get what they deserve in society is more divided. While a slim majority (56.5% combining agree and strongly agree) agree, a significant minority (43.6%) disagrees, showing that this is the most contentious item among respondents. There is general agreement that people are rewarded for talent and effort in society, but the consensus is weaker compared to school meritocracy. Deservingness in society is the most contentious, with significant disagreement among respondents.

*Market justice preferences in education, health and pensions by social and school meritocracy*


@fig-bivariate shows a series of graphs depicting the association between the variables of market justice preferences - in education, health, and pensions - and the variables of meritocratic perception at school (effort and talent) and in society (effort, talent, and deservingness) (see conceptual diagram in @fig-hypotheses). On the left, we observe the social meritocracy diagrams, while on the right the school meritocracy diagrams are shown. For the three variables of perception of meritocracy in society the relationship is clear, since the average of market justice preferences increases the more there is agreement that people are rewarded for their effort, merit, and talent. This relationship needs to be clarified in the case of the variables of perception of meritocracy at school. To the extent that there is more agreement that the perception of talent is essential for obtaining good grades, the average of market justice preferences increases, but this relationship is not as clear as with the variables of meritocracy in society. In addition, the graph does not show a clear trend in the relationship between the perception that effort is essential to obtain good grades and market justice preferences.

```{r echo=FALSE}
#| fig-cap: "Market justice preferences in education, health and pensions by social and school meritocracy"
#| fig-asp: 1.22
#| label: fig-bivariate

plot_soc_effort <- soc_effort %>% 
  mutate(justice = str_to_title(justice),
         justice = factor(justice, levels = c("Health", "Pensions", "Education"), labels = c("Health", "Pensions", "Education"))) %>% 
  arrange(esfuerzo_soc, justice) %>% 
  ggplot(aes(x = esfuerzo_soc, y = mean)) +
  geom_point(aes(color = justice, shape = justice), size = 2) +
  scale_y_continuous(labels = decimales, limits = c(1, 4)) +
  scale_x_continuous(limits = c(1, 4),
                     breaks = seq(1, 4, by = 1),
                     label = c("Strongly\n disagree", 
                               "Disagree", 
                               "Agree", 
                               "Strongly\n agree")) +
  scale_color_brewer(palette = "Paired") +
  labs(x = "Social effort",
       y = "Market justice preferences") +
  theme(legend.position = "none",
        text = element_text(size = 11))

plot_deservingness <- deservingness %>% 
   mutate(justice = str_to_title(justice),
         justice = factor(justice, levels = c("Health", "Pensions", "Education"), labels = c("Health", "Pensions", "Education"))) %>% 
  arrange(merito_soc, justice) %>% 
  ggplot(aes(x = merito_soc, y = mean)) +
  geom_point(aes(color = justice, shape = justice), size = 2) +
  scale_y_continuous(labels = decimales, limits = c(1, 4)) +
  scale_x_continuous(limits = c(1, 4),
                     breaks = seq(1, 4, by = 1),
                     label = c("Strongly\n disagree", 
                               "Disagree", 
                               "Agree", 
                               "Strongly\n agree")) +
  scale_color_brewer(palette = "Paired") +
  labs(x = "Deservingness",
       y = "Market justice preferences") +
  theme(legend.position = "none",
        text = element_text(size = 11))

plot_soc_talent <- soc_talent %>% 
   mutate(justice = str_to_title(justice),
         justice = factor(justice, levels = c("Health", "Pensions", "Education"), labels = c("Health", "Pensions", "Education"))) %>% 
  arrange(inteligencia_soc, justice) %>%  
  ggplot(aes(x = inteligencia_soc, y = mean)) +
  geom_point(aes(color = justice, shape = justice), size = 2) +
  scale_y_continuous(labels = decimales, limits = c(1, 4)) +
  scale_x_continuous(limits = c(1, 4),
                     breaks = seq(1, 4, by = 1),
                     label = c("Strongly\n disagree", 
                               "Disagree", 
                               "Agree", 
                               "Strongly\n agree")) +
  scale_color_brewer(palette = "Paired") +
  labs(x = "Social talent",
       y = "Market justice preferences") +
  theme(legend.position = "none",
        text = element_text(size = 11))

plot_school_effort <- school_effort %>% 
  mutate(justice = str_to_title(justice),
         justice = factor(justice, levels = c("Health", "Pensions", "Education"), labels = c("Health", "Pensions", "Education"))) %>% 
  arrange(esfuerzo_esc, justice) %>%  
  ggplot(aes(x = esfuerzo_esc, y = mean)) +
  geom_point(aes(color = justice, shape = justice), size = 2) +
  scale_y_continuous(labels = decimales, limits = c(1, 4)) +
  scale_x_continuous(limits = c(1, 4),
                     breaks = seq(1, 4, by = 1),
                     label = c("Strongly\n disagree", 
                               "Disagree", 
                               "Agree", 
                               "Strongly\n agree")) +
  scale_color_brewer(palette = "Paired") +
  labs(x = "School effort",
       y = "") +
  theme(legend.position = "none",
        text = element_text(size = 11))


plot_school_talent <- school_talent %>% 
  mutate(justice = str_to_title(justice),
         justice = factor(justice, levels = c("Health", "Pensions", "Education"), labels = c("Health", "Pensions", "Education"))) %>% 
  arrange(inteligencia_esc, justice) %>% 
  ggplot(aes(x = inteligencia_esc, y = mean)) +
  geom_point(aes(color = justice, shape = justice), size = 2) +
  scale_y_continuous(labels = decimales, limits = c(1, 4)) +
  scale_x_continuous(limits = c(1, 4),
                     breaks = seq(1, 4, by = 1),
                     label = c("Strongly\n disagree", 
                               "Disagree", 
                               "Agree", 
                               "Strongly\n agree")) +
  scale_color_brewer(palette = "Paired") +
  labs(x = "School talent",
       y = "",
       color = NULL,
       shape = NULL) +
  theme(text = element_text(size = 11))

leg <- get_legend(plot_school_talent)

plot_school_talent <- plot_school_talent + theme(legend.position = "none")

ggarrange(plot_soc_effort, plot_school_effort, plot_soc_talent,
          plot_school_talent, plot_deservingness, leg, ncol = 2, nrow = 3) +
  labs(caption = paste0("Source: Authors calculation based on First Study of Civic Education 2017",
                             " (N = ",dim(data_rec)[1],")"))
```

## Multivariate

This section presents the results of the multilevel models. We begin by showing the results of the cumulative link mixed models in @tbl-ordinal-reg for the ordinal dependent variables of justice in health, pensions, and education separately, to determine if the effect of the predictors differs across these items. Then, we present the results of the linear mixed-effects models for the average index of preference for market justice in @tbl-lineal-reg[^1]. For the health item, the intraclass correlation (ICC) obtained for the null model (not shown) indicates that 5% of the total variance of this item is associated with specific characteristics of the schools, while it is 3% for the pensions item and 4% for the education item (ICC based on [@hox_multilevel_2010, p.15]). For the average index of market justice preferences, the ICC reaches 4%, indicating that overall, only a small percentage of the total variance in students' market justice preferences is associated with school characteristics, thus limiting the possibility of finding effects at this second level of analysis.


[^1]: The tables shows the effects of individual and contextual-level independent variables related to our hypothesis; however, the models also incorporate the control variables utilized in the study for each of the three dependent variables separately and for the average index (complete models are available in the [Appendix](#appendix)).

```{r echo=FALSE, include=FALSE}

data_pension <- data_rec %>% 
    dplyr::select(-just_educ, -just_salud, -redistribucion)
data_pension <- na.omit(data_pension)
data_pension$just_pension_fact <- as_factor(data_pension$just_pension)

m0_p <- clmm(just_pension_fact ~ 1 + (1 | mrbd), data=data_pension)

performance::icc(m0_p, by_group = T) 

m1_p <- clmm(just_pension_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
    inteligencia_soc + esfuerzo_soc + merito_soc + (1 | mrbd), data=data_pension) 

m2_p <- clmm(just_pension_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
    inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + 
    libros_rec + acc_tec + internet_rec + (1 | mrbd), data=data_pension) 

m3_p <- clmm(just_pension_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
    inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + 
    libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + 
    cod_grupo + simce + (1 | mrbd), data=data_pension) 

modelos_pension <- list(m1_p, m2_p, m3_p)

# modelos_odd_pension <- map(tbls, odds_fun)

data_salud <- data_rec %>% 
    dplyr::select(-just_pension, -just_educ, -redistribucion)
data_salud <- na.omit(data_salud)
data_salud$just_salud_fact <- as_factor(data_salud$just_salud)

m0_h <- clmm(just_salud_fact ~ 1 + (1 | mrbd), data=data_salud)

performance::icc(m0_h, by_group = T) 

m1_h <- clmm(just_salud_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
    inteligencia_soc + esfuerzo_soc + merito_soc + (1 | mrbd), data=data_salud) 

m2_h <- clmm(just_salud_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
    inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + 
    libros_rec + acc_tec + internet_rec + (1 | mrbd), data=data_salud) 

m3_h <- clmm(just_salud_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
    inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + 
    libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + 
    cod_grupo + simce + (1 | mrbd), data=data_salud) 

modelos_health <- list(m1_h, m2_h, m3_h)

# modelos_odd_health <- map(tbls, odds_fun)


data_educ <- data_rec %>% 
    dplyr::select(-just_pension, -just_salud, -redistribucion)
data_educ <- na.omit(data_educ)
data_educ$just_educ_fact <- as_factor(data_educ$just_educ)

m0_e <- clmm(just_educ_fact ~ 1 + (1 | mrbd), data=data_educ)

performance::icc(m0_e, by_group = T) 

m1_e <- clmm(just_educ_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
    inteligencia_soc + esfuerzo_soc + merito_soc + (1 | mrbd), data=data_educ) 

m2_e <- clmm(just_educ_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
    inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + 
    libros_rec + acc_tec + internet_rec + (1 | mrbd), data=data_educ) 

m3_e <- clmm(just_educ_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
    inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + 
    libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + 
    cod_grupo + simce + (1 | mrbd), data=data_educ) 

modelos_educ <- list(m1_e, m2_e, m3_e)

#modelos_odd_educ <- map(tbls, odds_fun)

```

```{r echo=FALSE, results='asis'}
#| tbl-cap: "Cumulative link multilevel models of differential acces justificacion of health, pensions, and education"
#| label: tbl-ordinal-reg

ccoef <- list(
  inteligencia_esc = "School talent",
  esfuerzo_esc = "School effort",
  inteligencia_soc = "Social talent",
  esfuerzo_soc = "Social effort",
  merito_soc = "Deservingness",
  "educacion_recEd Media" = "Secondary",
  "educacion_recEd Tecnica" = "Higher tec.",
  "educacion_recUniversidad o posgrado" = "University or posgraduate",
  "educacion_recNs/Nr" = "Missing",
  "libros_recMás de 25" = "More than 25 books (Ref.= Less than 25)",
  "cod_grupoMedio bajo" = "SES Medium low",
  cod_grupoMedio = "SES Medium",
  "cod_grupoMedio alto" = "SES Medium high",
  cod_grupoAlto = "SES High",
  simceMedio = "Simce Medium",
  simceAlto = "Simce High"
)

texreg::htmlreg(l = list(modelos_health[[2]],
                         modelos_health[[3]],
                         modelos_pension[[2]],
                         modelos_pension[[3]],
                         modelos_educ[[2]],
                         modelos_educ[[3]]),   
                caption = NULL,
                stars = c(0.05, 0.01, 0.001),
                custom.note = "Note: Cells contain regression coefficients with standard errors in parentheses. Control variables are included. %stars",
                custom.header = list("Health" = 1:2, "Pensions" = 3:4, "Education" = 5:6),
                custom.model.names = rep(c("Model 1", "Model 2"), 3),
                custom.coef.map = ccoef,
                leading.zero = T,
                scalebox = 0.60,
                groups = list("Parental education (Ref.= 8th grade or less)" = 6:9,
                                "Socioeconomic level (Ref.= Low)" = 11:14,
                                "Achievement score (Ref.= Low)" = 15:16),
                custom.gof.rows = list("Controls"=rep("Yes",6)),
                include.loglik = FALSE, include.aic = FALSE)


```

Model 1 in @tbl-ordinal-reg for the three dependent variables of justice in differential access to health, pensions, and education demonstrates a similar trend regarding the effect of the perception of meritocracy on these items, with all effects being statistically significant (*p* \< 0.05). Regarding the variables of social meritocracy, the three variables---talent, effort, and deservingness---show that as these increase, the justification for differentiated access to pensions, education, and health also increases. In the context of school meritocracy, the effects are mixed: as perceived school talent increases, the justification for differentiated access to these services increases; conversely, as perceived school effort increases, the justification decreases, holding all other variables constant.

Additionally, Model 1 examines family status through parents' educational level and the number of books at home. For health, pensions, and education, a negative but statistically non-significant association with parents' education is observed. However, having more than 25 books at home shows a negative and statistically significant effect on market justice in health and pensions, suggesting less justification for differentiated access compared to those with fewer books. This effect is not seen in education.

Model 2 in @tbl-ordinal-reg incorporates school-level variables. The results show that high socioeconomic level schools show greater preferences for market justice in education compared to low socioeconomic level schools, with statistically significant results (*p* \< 0.05). This effect is not seen for health and pensions. Moreover, the findings suggest that schools with higher academic performance shows negative and statistically significant effects across all three items, suggesting these schools justify market allocation principles less. This pattern is also observed in medium-performing schools, but only for health and pensions.

```{r echo=FALSE, include=FALSE}

reg0_bienestar <- lmer(bienestar ~ 1 + (1 | mrbd), data = data_bienestar)

performance::icc(reg0_bienestar, by_group = T) 

reg1_bienestar <- lmer(bienestar ~ 1 + inteligencia_esc + esfuerzo_esc + 
                         inteligencia_soc + esfuerzo_soc + merito_soc + 
                         acc_tec + internet_rec +
                         (1 | mrbd), data=data_bienestar)

reg2_bienestar <- lmer(bienestar ~ 1 + inteligencia_esc + esfuerzo_esc + 
                         inteligencia_soc + esfuerzo_soc + merito_soc + 
                         educacion_rec + libros_rec + acc_tec + internet_rec +
                         (1 | mrbd), data=data_bienestar)

reg3_bienestar <- lmer(bienestar ~ 1 + mean_educ + cod_depe2 + cod_grupo +
                         simce + (1 | mrbd), data=data_bienestar)

reg4_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + 
                         inteligencia_soc + esfuerzo_esc + inteligencia_esc + 
                         educacion_rec + libros_rec + acc_tec + internet_rec + 
                         mean_educ + cod_depe2 + cod_grupo + 
                         (1 | mrbd), data=data_bienestar)

reg5_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + merito_soc + 
                         inteligencia_soc + esfuerzo_esc + inteligencia_esc + 
                         educacion_rec + libros_rec + acc_tec + internet_rec + 
                         mean_educ + cod_depe2 + cod_grupo + simce + 
                         (1 | mrbd), data=data_bienestar)


test01 <- anova(reg0_bienestar, reg1_bienestar, test = "Chisq")
test02 <- anova(reg0_bienestar, reg2_bienestar, test = "Chisq")
test03 <- anova(reg0_bienestar, reg3_bienestar, test = "Chisq")
test04 <- anova(reg0_bienestar, reg4_bienestar, test = "Chisq")
test05 <- anova(reg0_bienestar, reg5_bienestar, test = "Chisq")


test.pvalues1 <- test01$`Pr(>Chisq)`[2]
test.pvalues2 <- test02$`Pr(>Chisq)`[2]
test.pvalues3 <- test03$`Pr(>Chisq)`[2]
test.pvalues4 <- test04$`Pr(>Chisq)`[2]
test.pvalues5 <- test05$`Pr(>Chisq)`[2]

deviance1 <- test01$deviance[2]
deviance2 <- test02$deviance[2]
deviance3 <- test03$deviance[2]
deviance4 <- test04$deviance[2]
deviance5 <- test05$deviance[2]
```

```{r echo=FALSE, results='asis'}
#| tbl-cap: "Linear mixed-effects models for meritocracy perception and market justice preferences"
#| label: tbl-lineal-reg


ccoef <- list(
  "(Intercept)" = "Constant",
  inteligencia_esc = "School talent",
  esfuerzo_esc = "School effort",
  inteligencia_soc = "Social talent",
  esfuerzo_soc = "Social effort",
  merito_soc = "Deservingness",
  "educacion_recEd Media" = "Secondary",
  "educacion_recEd Tecnica" = "Higher tec.",
  "educacion_recUniversidad o posgrado" = "University or posgraduate",
  "educacion_recNs/Nr" = "Missing",
  "libros_recMás de 25" = "More than 25 books (Ref.= Less than 25)",
  "cod_grupoMedio bajo" = "SES Medium low",
  cod_grupoMedio = "SES Medium",
  "cod_grupoMedio alto" = "SES Medium high",
  cod_grupoAlto = "SES High",
  simceMedio = "Simce Medium",
  simceAlto = "Simce High"
)


texreg::htmlreg(list(reg1_bienestar, reg2_bienestar, reg4_bienestar, reg5_bienestar),
                custom.model.names = c("Model 1",
                                       "Model 2",
                                       "Model 3",
                                       "Model 4"),
                caption = NULL,
                stars = c(0.05, 0.01, 0.001),
                custom.note = "Note: Cells contain regression coefficients with standard errors in parentheses. Control variables are included. %stars",
                custom.coef.map = ccoef,
                leading.zero = T,
                scalebox = 0.60,
                groups = list("Parental education (Ref.= 8th grade or less)" = 7:10,
                              "Socioeconomic level (Ref.= Low)" = 12:15,
                              "Achievement score (Ref.= Low)" = 16:17),
                custom.gof.rows = list("Controls"=rep("Yes",4),
                                       "Deviance" = c(deviance1,
                                                    deviance2,
                                                    deviance4,
                                                    deviance5),
                                       "Deviance Test (p)" = c(test.pvalues1,
                                                             test.pvalues2,
                                                             test.pvalues4,
                                                             test.pvalues5)),
                include.loglik = FALSE,include.aic = FALSE)

```

@tbl-lineal-reg presents the results of linear mixed-effects models for the market justice preference index. Model 1 shows that higher perceived societal meritocracy, through effort, talent, and deservingness, is positively and significantly associated with market justice preferences. School meritocracy perceptions yield mixed effects: perceived talent increases market justice preference, while perceived effort decreases it, both statistically significant (*p* \< 0.001). Model 2 includes family status, finding no statistically significant effects from parents' education or the number of books at home, both showing negative directions.

Regarding school characteristics, Models 3 and 4 introduce socioeconomic level and academic performance. The results of Model 3 suggest no statistically significant relationship between school socioeconomic levels and market justice preferences. However, Model 4 indicates that schools with medium ($\beta$ = -0.10) and high ($\beta$ = -0.24) academic performance have lower market justice preferences compared to low-performing schools, with these effects being statistically significant when controlling for other variables.

```{r echo=FALSE}
#| tbl-cap: "Interactions effects"
#| label: tbl-interact

pacman::p_load(kableExtra)
load("input/data/interact.RData")

collapse_rows[c(-4),] %>% 
  as_tibble() %>% 
  kbl(.,
      align = "lccccc", booktabs = T,
      col.names =  c("Variable", 
                     "School talent", 
                     "School effort",
                     "Social talent",
                     "Social effort", 
                     "Deservingness")) %>% 
  add_header_above(c(" " = 1,
                     "Market Justice Preferences"=5), 
                   bold = T) %>% 
  kable_styling(latex_options = c("scale_down","HOLD_position"),
                full_width = F) %>% 
  column_spec(1, bold = T, width = "4cm") %>%
  column_spec(column = 2:6,width = "1.5cm") %>%
  collapse_rows(columns = 1:2, valign = "middle") %>% 
  kableExtra::add_footnote(label = "Note: ***p < 0.001, **p < 0.01, *p < 0.05", 
                           notation = "none")


```

@tbl-interact displays the interaction terms related to our hypothesis. The results suggest that students' family background moderates the relationship between their meritocratic perceptions and their justification of market allocation in social services. Specifically, the relationship between social effort and market justice preferences is less positive for students whose parents have university or postgraduate education. Lower-status students (measured by parental education) prefer more market justice when they strongly adhere to meritocratic perceptions of deservingness. A similar moderating effect is seen with family cultural capital: the relationship between effort and market justice preferences is less positive for students with more than 25 books at home.

At the school level, we observe that school status and students' meritocratic beliefs moderate their preferences for market justice. High-status schools show less preference for market justice when their students, on average, have a greater perception of meritocracy, as measured by effort, deservingness, and talent. Conversely, low-achieving schools prefer more market justice when their students have stronger meritocratic beliefs. @fig-interaction illustrates this moderating effect of school achievement on the relationship between deservingness and market justice. No moderating effects were observed at the school level for students' meritocratic beliefs regarding effort or talent in achieving good grades.

```{r echo=FALSE, include=FALSE}
load("input/data/data_rec.RData")
data_rec <- data_rec %>% rowwise() %>%  mutate(educacion_rec2 = case_when(educacion_rec=="Ed Basica"~"Low level",
                                                 educacion_rec=="Ed Media"~"Low level",
                                                 educacion_rec=="Ed Tecnica"~"Low level",
                                                 educacion_rec=="Universidad o posgrado"~"University or Postgraduate",
                                                 educacion_rec=="Ns/Nr"~"Missing"
                                                 ))
data_rec$educacion_rec2 <- factor(data_rec$educacion_rec2, levels = c("Low level", "University or Postgraduate", "Missing"))


data_rec <- data_rec %>% rowwise() %>%  mutate(cod_grupo2  = case_when(cod_grupo=="Bajo"~"Bajo",
                                                                       cod_grupo=="Medio bajo"~"Bajo",
                                                                       cod_grupo=="Medio"~"Bajo",
                                                                       cod_grupo=="Medio alto"~"Alto",
                                                                       cod_grupo=="Alto"~"Alto"))
data_rec$cod_grupo2 <- factor(data_rec$cod_grupo2, levels = c("Bajo", "Alto"))
data_rec$simce <- factor(data_rec$simce, labels = c("Low", "Medium", "High"))


data_bienestar <- na.omit(data_rec)
#data_bienestar <- data_bienestar %>% mutate(pension_rec = 5-just_pension,
#                                            educ_rec = 5-just_educ,
#                                            salud_rec = 5-just_salud)
data_bienestar <- data_bienestar %>% rowwise() %>% dplyr::mutate(bienestar = mean(c(just_educ, just_salud, just_pension), na.omit=TRUE))
data_bienestar$bienestar <- round(data_bienestar$bienestar, 2)

reg26.2_bienestar <- lmer(bienestar ~ 1 + esfuerzo_soc + inteligencia_soc + esfuerzo_esc + inteligencia_esc + educacion_rec2 + libros_rec + acc_tec + mean_educ + cod_depe2 + cod_grupo2 + merito_soc*simce + internet_rec + (1 + simce | mrbd), data=data_bienestar)

plot_interact<- plot_model(reg26.2_bienestar, type = "int", mdrt.values="minmax", title = "", colors = c("red", "white", "blue")) +
  ylim(1,4)+
  labs(x = "Deservingness", y="Market Justice Preferences", colour = "School Simce")+
  theme_bw()
```

```{r echo=FALSE, warning=FALSE}
#| fig-cap: "Interaction between deservingness and justification of inequality by school SIMCE achievement"
#| label: fig-interaction

plot_interact

```
