---
title: "market-justice-preferences"
author: "Equipo EDUMER"
format: html
editor: visual
bibliography: "input/bib/just-ineq-merit.bib"
---

# The socialization of meritocracy and market justice preferences at school

## Introduction

Since its origins, educational institutions have been related to the idea of social mobility and access to better opportunities. Therefore, the consistent evidence of the high level of social reproduction at the school level represents a threat to the promise of education and a meritocratic system [@bourdieu_reproduction_1990]. A large part of the research in this field at an international level has addressed the extent to which the social origin of students affects their academic results and their life opportunities [@vonhippel_test_2019], confirming that schools have severe difficulties in closing the gaps of origin. Besides this socioeconomic perspective on school opportunities, recent research has addressed to what extent inequalities in the school context are also influencing students' perceptions, beliefs, and attitudes: Are social inequalities perceived at the school context? Are they rejected by the students, particularly those who are worst-off in socioeconomic terms? Or, Is there evidence at the school level that social inequalities are tolerated and even justified? [@batruch_belief_2022; @wiederkehr_belief_2015]. 

In the present paper, we deal with the justification of social inequalities by eighth-grade students in Chile, a country characterized by a highly stratified educational system. In particular, we focus on market justice preferences (Lane, 1986), which refer to the preferences for distributing public goods (as health and education) based on criteria such as competition and payment capacity. Although from a rational point of view it could be expected an opposition to market justice by the underprivileged majority,  we argue that in a social environment characterized by a promotion of meritocratic ideals - as the school system - would lead to the opposite: a larger market justice preferences. 

Given that the school environment has an important focus on performance, achievement and acknowledgment, meritocracy has been one of the principal concepts used for understanding and even for justifying performance differences among students. Meritocracy is a distributive system based on the belief that people should be rewarded and promoted based on their abilities, knowledge, and achievements (Young, 1958). It is often seen as a way to create equal opportunities and fairness, as individuals can rise to positions of power and influence based on their own merit rather than their background or connections. However, some argue that meritocracy can actually lead to tolerating or even justifying social inequalities, as it can create a hierarchy where those who already have resources and advantages are more likely to succeed. In this regard, a great deal of academic research about meritocracy delves into the assessment of to what extent rewards and privileges in society are related to merit, emphasizing the so-called unfulfillable promise of meritocracy (Mijs, 2016). Complementing this agenda, a second and emerging research area deals with subjective aspects of meritocracy, such as perceptions and beliefs. 

\[definición de market justice, welfare marketization, (Lindh)\]

\[contexto Chileno market justice, dar cifras/porcentajes de privatización de salud, educación y pensiones; asociar al tema de la calidad (lo público es peor)\]

The perception of meritocracy refers to how individuals view and understand the concept of meritocracy in their own society (Duru-Bellat & Tenret, 2012; Castillo et al., 2019). This perception can vary greatly depending on individual experiences, social, economic, and cultural background. Some people may see meritocracy as a fair and just system that allows anyone to succeed based on their abilities and hard work. In contrast, others may view it as a myth or a cover for existing power dynamics and inequality, serving to maintain and even reinforce inequality (Lampert, Mijs, etc). Based in this last perspective, we argue that individuals with a higher perception of meritocracy will show a larger justification of social inequalities, as individual achievement would be seen as rewarded and social policies as less necessary (Batruch et al, 2021).

Most of the research that has related meritocratic beliefs to inequality justification so far has only considered adults, leaving aside the study of how beliefs in this field develop at student age as well as the impact of the school context and the family as the main socialization agencies. Regarding schools, the way in which they deal with unequal conditions of origin has been linked to the \_hidden curriculum\_ (Chafel, 1997, https://doi.org/10.1002/J.2379-3988.1997.TB00004.X), whereby students learn about distributive norms in society and mechanisms of justification of social differences. Based on recent studies that relate school meritocracy to the justification of economic inequalities in the adult population \[\@batruch_belief_2022;\@wiederkehr_belief_2015\], the central hypothesis guiding this research is that school-age students with a higher perception of meritocracy - both at school and at the societal level - will show a larger justification of social inequalities, as individual achievement would be seen as appropriately rewarded and social mechanisms for correcting inequalities as less necessary (Batruch et al, 2021).

\~

\[introducir distinción sobre percepción de meritocracia a nivel escolar y a nivel de la sociedad\]

\[resaltar 3 focos de innovación: market justice preferences at school level, relación con meritocracia, y esquema de meritocracia en la escuela y en la sociedad\]

The present paper deals with the association between the perception of meritocracy and the justification of social inequalities, with two main focuses. Firstly, it assesses the justification of social inequalities in social policy domains, such as health, pensions, and education. We argue that individuals who perceive more meritocracy would be more willing to justify better services in these domains for those with higher incomes.  Secondly, we focus on the student-age population as we point out that it is possible to track down the origin of meritocratic beliefs (and their consequences) to early socialization processes. To this regard, we take into account the family and the school as two main socialization agencies that play a significant role in the socialization of cultural beliefs by transmitting cultural norms, values, and expectations to young people. 

## Justification of inequality

Research on social stratification beliefs, which explore individual perceptions of who deserves what and why (Kluegel & Smith, 1987), highlights that people’s explanations and justifications of social inequality are closely tied to their judgments of deservingness. The influence of ideologies (Wegener & Liebig, 1995) and cultural schemas (Homan et al., 2017) is pivotal in shaping these explanations by offering symbolic representations that frame societal structures and expectations. While significant attention has been paid to wage inequality, income distribution, and payment differentials in the literature (Castillo, 2011; Evans et al., 2010; Jasso, 1999; Shariff et al., 2016),  there has been less examination of public beliefs about which life domains should be governed by market relations and even less about children’s acceptance or rejection of these market principles. This oversight is notable given the extensive encroachment of market logic into public goods, welfare policy, and social services over the past five decades (Centeno & Cohen, 2012; Harvey, 2007), affecting areas such as pensions, health services, and education. The justification of social inequality based on marketype criteria has been conceptualized as the individuals’ adherence to one specific justice evaluation, the market justice, that is, affording legitimacy to the allocation of goods and services based on prices and individuals’ ability to pay (Boltanski & Chiapello- 2007; Lane 1986; Streeck 2012).

Robert E. Lane proposed the underpinnings of market justice, which he differentiated from political justice. For him, “it is the genius of the market to stimulate wants without at the same time stimulating a sense of deserving more than one gets” (1986, p. 384). Following the theory of relative deprivation –a social phenomenon arising when individuals cannot afford what most others in their environment can (Merton 1950)– Lane notes that, in market settings, social comparisons are more likely to motivate increased effort rather than feelings of acute injustice because individuals attribute outcomes to their actions. Although empirical research has shown that, contrary to Lane’s observation, relative deprivation, even in market settings, produces feelings of dissatisfaction, anger, and resentment that might motivate forms of collective action such as protests and revolt (Greitemeyer and Sagioglou 2016; Mishra and Carleton 2015; Smith et al. 2012; Power 2018; Power, Madsen, and Morton 2020), the conceptualization of market justice has been since then closely coupled to the merit principle for allocating outcomes: the claim that the unequal levels of well-being individuals enjoy ought to be, to some extent, a function of their talents and efforts, regardless of their needs or membership, the two latter being the realm of political justice and its closely coupled principles of need –allocating outcomes to those who require them most– and equality –allocating the same outcome to everyone–, which have been at the center of welfarism (see Wilson 2003). The underlying notion of market justice also resonates in its application to welfare regimes, providing a framework for understanding the varied global approaches to managing social services.

The management of social services manifests in varying approaches across nations, with substantial differences in funding and delivery methods (Jensen, 2008; Stoy, 2014). Nordic countries, for example, predominantly employ public agencies to produce and provide social services, funding these through collective taxation and offering them in kind to the majority of citizens. This system prioritizes political justice, placing it above market mechanisms in accessing services. In contrast, other countries rely more heavily on for-profit entities and private funding, where service distribution depends mainly on individual financial capacity of paying user fees, highlighting the influence of market justice in service allocation. The trend toward marketization of welfare services has been growing since the 1980s (Salamon, 1993), and this shift is increasingly evident even in countries where market solutions have traditionally had a minor role in social policy (Sivesind, 2017).

The question arises whether adults and children justify unequal access to welfare services based on market justice principles. Influenced by theories of policy feedback, which suggest that social welfare policies can reinforce (positive feedback) or undermine (negative feedback) previous policy trajectories (Fernandez and Jaime-Castillo 2013; Pierson 2000; Weaver 2010), citizens’ beliefs about market justice are likely also shaped by the institutional and social contexts they encounter. Indeed, the justification of inequality in access to essential services like education and health, based on one’s ability to pay, shows significant variation across countries, as demonstrated by international surveys, although they do not usually include children in their samples. For instance, the 2019 International Social Survey Program (ISSP) provides insights into how adults perceive inequality in accessing welfare services. Figure 1 illustrates the variation in agreement levels regarding whether it is just for individuals with higher incomes to purchase better education or healthcare. Notably, in 18 of the 29 countries analyzed, there is a greater justification for market inequality in healthcare access than in education. Nevertheless, the general sentiment typically ranges from 'somewhat unjust' to 'neither just nor unjust, with mixed feelings.

```{r}
#| fig-cap: "Average of market justice preferences by country"
```

While contemporary policy developments have increasingly embraced the marketization of social welfare in areas such as education, pensions, and health services, questions about public acceptance of these changes persist. Lindh’s (2015) analysis of ISSP 2009 data from 17 OECD countries reveals a general lack of support for market-based distribution of social services, suggesting widespread disapproval of market stratification of essential services. This finding is corroborated by Soler-Martínez and colleagues’ (2023) research from Latinobarómetro 2020 across 18 Latin American countries, where concerns about health and education access predominated over income inequality. These results indicate that reforms toward welfare marketization are typically driven by elite political decisions rather than grassroots demand. 

Despite high-income inequality and limited social mobility in Latin America, there is a prevalent belief that individuals are solely responsible for their economic outcomes, a view that varies across the region (Bucca, 2016; Chong et al., 2008; Torche, 2014; Salgado and Castillo, 2023). The reliance on private welfare providers and widespread user fees (Molyneux, 2008) adds complexity to this context. Yet, research on children’s justification of market-based inequalities in accessing welfare services remains limited, especially in Latin America, highlighting a significant gap in understanding how younger generations view market-based access to welfare and whether these views are associated with their meritocratic beliefs. 

## Meritocracy

The concept of meritocracy frequently appears nowadays when analyzing cultural determinants of social inequalities. In general, it is mentioned as a value associated with justice, as it would link efforts and talents with rewards in an equitable manner. This normative sense is quite far from its original formulation by Michael Young (1958) in the satirical novel "The Rise of Meritocracy", where it ironically represented a mechanism for reproducing the inequalities of origin. The meritocratic ideal had remained relatively unchallenged until a series of recent publications turned into its potential consequences for maintaining social inequality. Perhaps one of the most recent sources in this line is Michael Sandel's "The Tyranny of Merit", where he strongly questions the implications of carrying out the principle of merit in societies that do not guarantee equal opportunities and that generate a feeling of scarce recognition and appreciation of those who receive lesser rewards: "In society's eyes, and perhaps also their own, their work no longer signified a valued contribution to the common good." (Sandel, pp).

Empirical research on meritocracy has increased along with the philosophical-normative discussion on meritocracy in recent years. Particularly from a sociological perspective, meritocracy has been used in research on social mobility to characterize societies with low mobility that threaten the meritocratic ideal (Goldthorpe). More recently, sociology and social psychology research has attended to the subjective aspects vis-a-vis beliefs in meritocracy. The label of beliefs in this realm covers a series of areas, such as attitudes, perceptions, and preferences (Castillo et al), whereby most of the link this subjective dimensions to individual socio-structural factors and  context-level determinants. For instance, some studies have analyzed how those with greater privileges believe more in meritocracy (Reynolds & Chan 2014), how greater economic inequality increases meritocratic beliefs (Mijs 2021), and how larger inequality affects meritocratic beliefs (Morris et al 2022). Based on these findings, a research agenda has been reinforced on the legitimizing role of meritocracy, in line with previous studies using the concept of a just world (Lerner, Dalbert) and the theory of system justification (Jost & Major).

How do meritocratic beliefs legitimize inequalities? Empirical studies have used experiments and surveys to address this question. For instance, the evidence suggests that just world beliefs correlate negatively with support for redistributive compensation systems (Frank et al., 2015). Conversely, individuals tend to support redistribution when they believe that the disadvantaged lack the opportunities to succeed (Evans & Kelley, 2018). Almås and colleagues (2020) found that in a relatively unequal society (the United States), the highly educated accept inequality significantly more than the less educated because they perceive inequality as justifiable owing to differences in productivity (i.e., merit), whereas in a relatively equal society (Norway), the less educated accept inequality more, but not significantly more than the highly educated because meritocratic values are less prevalent. Barr and Miller (2020) also addressed this triple interaction between the level of inequality in a society, the individual level of education, and the perceived origin of the disparity (either by luck or effort) to determine the extent to which inequality is accepted. They found that the interaction's mechanism varies depending on the compared societies. Finally, García-Sánchez and colleagues (2020), using data for 41 countries from the International Social Survey Programme (ISSP), found that the perceived size of the income gap correlated positively with support for progressive taxation. Still, this association was weaker among those who endorsed meritocratic and equal opportunity beliefs. In the same line, experimental research by Durante & Putterman (2009) subjects support less redistribution when the initial distribution is determined according to task performance.

Research about meritocratic beliefs at school age is rather scarce, leaving a wide research gap as schools are one of the primary socialization institutions where achievement based on merit explains success (Erivwo et al 2021). Nevertheless, it is possible to find some initial works focused on the area of distributive justice at school that are closely related to meritocratic beliefs, as the ones by Resh and Sabbagh (quotes). Using justice in grade obtention as a measure of distributive justice and meritocracy (Sabbagh et al 2006), they find for instance that a larger sense of distributive justice about grades is associated to higher socio-economic status (Resh 2010), have a positive effect on liberal democratic orientation and on trust in people and in formal institutions (Resh & Sabbagh 2013), and tend to refrain from violence and to engage to a greater extent in extra-curricular school activity and community volunteering (Resh & Sabbagh 2017).

## Children's judgments of inequality, schools, and family background

Research indicates various socialization practices at families and schools during childhood and adolescence that impact dispositional and behavioral tendencies concerning justifications of inequality in adult life. The differences in economic understanding across age groups are consistent with research on cognitive development (Choudhury et al., 2006). Adolescents with mature socio-cognitive abilities tend to express stronger preferences for fairness than infants and children (Wynn et al., 2017). As children grow older, they become more likely to behave fairly, with their early-emerging strict egalitarianism being replaced by an increasing endorsement of fairness principles and engagement in collaborative activities (Huppert et al., 2019; McAuliffe et al., 2017). In these activities, their fairness views consider individual contributions, merits, and circumstances (Almås et al., 2010; Huppert et al., 2019; Sigelman & Waitzman, 1991). Engelmann and Tomasello (2019) claim that children’s sense of fairness emerges at three years old, and we can observe it in collaborative activities, where they accept inequality if the procedure gives everyone an equal chance. Therefore, children at this age respond to unequal distributions based on interpersonal concerns, as they already demand equal respect. In any case, between 3 and 8 years of age, inequitable and anti-meritorious allocations are evaluated more negatively, but equitable and meritorious allocations are not evaluated more positively (Elenbaas, 2019).

Some research shows that the social environment in which children develop, such as family and school, is associated with their prosocial behaviors by playing an essential role in the transmission of equity norms (Schunk & Zipperle, 2023; Kosse et al., 2020). In fact, schools contribute to institutionalizing and reproducing inequality by promoting values, norms, practices, and languages familiar to higher-class families because the dominant group’s culture shapes educational institutions (Bourdieu & Passeron, 1990). Middle- and upper-class students are better equipped to face academic challenges and are more familiar with academic expectations (Mikus et al., 2019). Such familiarity represents cultural capital in educational contexts because higher-status students come to school ready to meet these expectations and reap the benefits (Jack, 2016; Khan, 2011). Conversely, lower-status children lacking cultural capital must catch up while experiencing inequitable comparisons (Goudeau & Croizet, 2017). Additionally, academic achievement is treated as the outcome of dispositional factors (e.g., pupils’ efforts and talents or lack of them) rather than the result of differential access to critical resources. Due to the meritocratic frame schools encourage, both low- and high-status individuals believe that students’ success or failure is not due to their family background but rather to differences in efforts and talents (Darnon et al., 2018). In this sense, we believe that the perception of meritocracy can influence students' judgments about market justice preferences. Furthermore, we believe there is a difference between students' perceptions of meritocracy based on their own experience in school and what they perceive in society at large. Consistently, our first two hypotheses are:

H1a: Students who perceive that there is more meritocracy at school will show larger market justice preferences

H1b: Students who perceive that there is more meritocracy in society will show larger market justice preferences

Family background and family socialization practices also contribute to children’s and adolescent’s market justice preferences. For example, Almås and colleagues (2017) found that adolescents from low-socioeconomic-status families are likelier to have an egalitarian fairness view and consider an equal distribution as fair in a situation with unequal merits. The authors speculate that differences in socialization practices across status groups might bring about, to a great extent, the fairness views of children and adolescents because social status seems to interact with these evaluations (e.g., Hvidberg et al., 2023). 

The classic work of Kohn showed that middle-class parents value the expression of internal states and emotions, such as self-control, curiosity, happiness, and consideration, while working-class parents promote deference, obedience, and conformity to authority (Kohn, 1963; Kohn & Schooler, 1969). Although parents from all social backgrounds encourage individualism in their children, this shared norm translates into different forms in high and low social classes (1999). Acemoglu (2021) claimed that the values families impart to their children interact with social mobility. Because obedience is a valuable characteristic for employers, in low-wage and social mobility environments, low-income families impart values of obedience to their children to prevent disadvantaging them in labor markets. On the one hand, children from privileged families are socialized to adopt a clear conception of individualism that highlights their internal states, independence, and idiosyncrasies. In contrast, children from disadvantaged families are socialized to support a more balanced view of individualism that considers personal characteristics as resources to overcome collective impediments on the path to upward mobility (Iacoviello & Lorenzi-Cioldi, 2019). In this way, we believe that there are differences in the socialization of values according to socioeconomic differences that could influence market justice preferences and, therefore, we propose the following hypothesis:

H2: Students from families of higher social status will show larger market justice preferences.

Recent empirical research has demonstrated that the institutional design of schools, coupled with the meritocratic ideology it fosters, significantly influences children’s and adolescents’ views on inequality and deservingness. For example, Jonsson and Beach’s (2015) study revealed that higher-status adolescents in Sweden tend to perpetuate social class stereotypes while describing the vocational and academic tracks. Academic track students are depicted as wealthy, intelligent, ambitious, and diligent, while vocational track students are characterized as poor, unambitious, unintelligent, and lackadaisical. These stereotypes help individuals maintain a sense of superiority over others and legitimize the prevailing social hierarchies and economic disparities (Jost & Burgess, 2000).

H3: Students from schools of higher social status will show larger market justice preferences.

H4: Students from schools with higher average levels of academic achievement will show larger market justice preferences.

\[interacciones\]

H5: The perception of meritocracy in school and society will moderate the effect of family social status on market justice preferences.

H6: The perception of meritocracy in school and society will moderate the effect of school status on market justice preferences.

H7: The perception of meritocracy in school and society will moderate the effect of school academic achievement on market justice preferences.

### Summary of hypotheses

```{r}
#| fig-cap: "Summary of hypotheses"
```

## Methods

### Data

The main data source for the analysis is the First Study of Citizenship Education in Chile, carried out by the Education Quality Agency of the Ministry of Education. The application date was November 9, 2017. The target population of this study is Eighth-grade students from 242 schools. In the data, there are 8,589 students and 6,770 parents. This database was analyzed with the R package “ResponsePatterns” to detect possible repetitive and “careless” response patterns and thus contribute to a higher quality of research data (Gottfried, et al 2022). Responses from 171 students and 79 parents were removed, which, when merged, gave a total of 6,270 valid cases.

The analysis of school variables includes data from the Ministry of Education's SIMCE 2017 database. This database contains information at the school level, such as the administrative dependency, its socioeconomic classification, and the achievement scores obtained in the mathematics and language census tests. It is available for free use on the MINEDUC \[web page\].

After eliminating missing cases, the final sample used in the analysis was based on 5,047 students and parents of 231 schools for the dependent variable of access to social services.

### Variables

**Dependent variables**

This study has three dependent variables related to the justification of social inequality in specific policy domains. The first asks whether access to social services should be conditional on income, i,.e., “It is just that in Chile people who can pay have a better education for their children”. Students rated their preferences using the following responses: “strongly disagree”, “Disagree”, “Agree”, and ”strongly agree”. An average index is built with these items (Cronbach's Alpha = 0,86). (Apéndice: items en español y su correspondiente traducción al inglés)

Table 1 shows the items used, their response categories, and their frequencies.

```{r}
#| tbl-cap: "Dependent variables"
```

**Independent variables**

For the primary independent variable, the perception of meritocracy, five items address the perception of rewards according to effort and intelligence at the school and societal levels. At the school level, students answer whether “Intelligence is important to get good grades” and “Effort is important to get good grades”. At the societal level, students respond to the following questions: “In Chile, people are rewarded for their efforts”, “In Chile, people get what they deserve”, and “In Chile, people are rewarded for their intelligence and skills”. Each item was answered on a four-point scale ranging from “Strongly disagree” to “Strongly agree”.

The rest of the independent variables are divided into individual and school levels. At the individual level, family socioeconomic status was measured by the parents' highest educational level and the number of books at home. Likewise, an index of access to technology includes the number of computers, tablets, and cell phones at home, as well as whether there is an Internet connection. Table 2 shows the items used, their response categories, and their frequency.

```{r}
#| tbl-cap: "Independent variables"
```

The school-level variables are the administrative dependency of the school, the socioeconomic classification made by the Ministry of Education, the level of performance in the SIMCE test of the school, and the proportion of parents with university or postgraduate degrees. Table 3 shows the items used, response categories, and frequency.

```{r}
#| tbl-cap: "School context variables"
```

### Methods

The data has a hierarchical structure of students nested in schools, so the model estimation is performed in a multilevel (random effects) framework. This modeling approach lets us correctly estimate individual and contextual effects in a single model. We estimate cumulative link mixed models for the ordinal dependent variables, whereas we use linear mixed effects models for the average index of inequality justification.

The hypotheses of this research were pre-registered in the Open Science Framework platform of the Center for Open ScorrectlyOSF), the access to the document is available at this \[link\](https://doi.org/10.17605/OSF.IO/UFSDV). The statistical analysis of this research was performed using the free software R version 4.1.3.

## Analysis

### Descriptive analysis

```{r}
#| fig-cap: "Market justice preferences in education, health and pensions by social and school meritocracy"
```

Figure 3 shows a series of graphs depicting the association between the variables of market justice preferences - in education, health, and pensions - and the variables of meritocratic perception at school (effort and talent) and in society (effort, talent, and deservingness) (see conceptual diagram en figure X). On the left, we observe the social meritocracy diagrams, while on the right the school meritocracy diagrams are shown. For the three variables of perception of meritocracy in society the relationship is clear, since the average of market justice preferences increases the more there is agreement that people are rewarded for their effort, merit, and talent. This relationship needs to be clarified in the case of the variables of perception of meritocracy at school. To the extent that there is more agreement that the perception of talent is essential for obtaining good grades, the average of market justice preferences increases, but this relationship is not as clear as with the variables of meritocracy in society. In addition, the graph does not show a clear trend in the relationship between the perception that effort is essential to obtain good grades and market justice preferences.
